---
title: "ESM_421_LCV_life"
author: "Ruben Sanchez Ramirez"
date: "1/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(ggplot2)
library(plyr)
```

```{r}
candidate <- c("Bernie Sanders", "Elizabth Warren", "Joe Biden", "Trump")
ideology <- c(10.5, 28, 44.5, 67.75)

candidate_ideology <- data.frame(candidate, ideology)

ggplot(data = candidate_ideology)+
  geom_bar(aes(x = ideology, y =.5 ),
           stat = "identity",
           width = 0.3)+
  geom_vline(aes(xintercept = 1),
             color = "black", linetype="solid")+
  geom_vline(aes(xintercept = 50),
             color = "black", linetype="solid")+
  geom_vline(aes(xintercept = 100),
             color = "black", linetype="solid")+
  geom_hline(yintercept=0.25)+
  xlim(-0.5,100.5)+
  scale_x_continuous(breaks = c(1,50,100),
                       labels = c(1,50,100))+
  theme_classic()
```


Read the "2018_LCV_Scores_P_set.csv" data set for question 4 :
```{r}
LCV_life <- read_csv("2018_LCV_Scores_P_Set.csv") %>% 
  janitor::clean_names()
```

Create sub-sections: for the House and the Senate:
```{r}
house_life <- LCV_life %>% 
  filter(chamber == 1)

senate_life <- LCV_life %>% 
  filter(chamber == 2)
```

Now, lets start by finding the median LCV scores and names in both the House and the Senate:
```{r}
house_LCV <- as.vector(house_life$lifetime_score)

house_median <- median(house_LCV)
house_closest <- which(abs(house_LCV-house_median)==min(abs(house_LCV-house_median)))
house_median_LCV <- house_life$lifetime_score[house_closest]
house_median_name <- house_life$name[house_closest]
house_median_LCV
house_median_name

senate_LCV <- as.vector(senate_life$lifetime_score)

senate_median <- median(senate_LCV)
senate_closest <- which(abs(senate_LCV-senate_median)==min(abs(senate_LCV-senate_median)))
senate_median_LCV <- senate_life$lifetime_score[senate_closest]
senate_median_name <- senate_life$name[senate_closest]
senate_median_LCV
senate_median_name

```

Now, lets find the veto override (2/3) score and names in both the House and the Senate, for a President with LCV = 0:
```{r}
senate_v_piv <- ceiling(sum(sapply(senate_LCV,length)) * (2 / 3))

senate_veto_0 <- senate_LCV %>%
  sort() %>% 
  tail(senate_v_piv) %>% 
  head(1)
senate_v0_LCV <- senate_life$lifetime_score[senate_LCV==senate_veto_0]
senate_v0_name <- senate_life$name[senate_LCV==senate_veto_0]
senate_v0_LCV
senate_v0_name

house_v_piv <- ceiling(sum(sapply(house_LCV, length)) * (2 / 3))

house_veto_0 <- house_LCV %>% 
  sort() %>% 
  tail(house_v_piv) %>% 
  head(1)
house_v0_LCV <- house_life$lifetime_score[house_LCV==house_veto_0]
house_v0_name <- house_life$name[house_LCV==house_veto_0]
house_v0_LCV
house_v0_name

```

Lets complete this senario by finding the filibuster pivot (3/5 - 60% | Right and Left) scores and names, for the Senate (NO filibuster in the House):
```{r}
senate_f_piv <- ceiling(sum(sapply(senate_LCV,length)) * (3 / 5))

senate_fili_R <- senate_LCV %>%
  sort() %>% 
  tail(senate_f_piv) %>% 
  head(1)
senate_fR_LCV <- senate_life$lifetime_score[senate_LCV==senate_fili_R]
senate_fR_name <- senate_life$name[senate_LCV==senate_fili_R]
senate_fR_LCV
senate_fR_name

senate_fili_L <- senate_LCV %>% 
  sort() %>% 
  head(senate_f_piv) %>% 
  tail(1)
senate_fL_LCV <- senate_life$lifetime_score[senate_LCV==senate_fili_L]
senate_fL_name <- senate_life$name[senate_LCV==senate_fili_L]
senate_fL_LCV
senate_fL_name

```

Now, lets see how things would change in a senario with a President of an LCV = 100 (Nothing else changes). The only difference would be in the veto override (2/3) scores and names in both the House and the Senate:
```{r}
senate_veto_100 <- senate_LCV %>%
  sort() %>% 
  head(senate_v_piv) %>% 
  tail(1)
senate_v100_LCV <- senate_life$lifetime_score[senate_LCV==senate_veto_100]
senate_v100_name <- senate_life$name[senate_LCV==senate_veto_100]
senate_v100_LCV
senate_v100_name

house_veto_100 <- house_LCV %>% 
  sort() %>% 
  head(house_v_piv) %>% 
  tail(1)
house_v100_LCV <- house_life$lifetime_score[house_LCV==house_veto_100]
house_v100_name <- house_life$name[house_LCV==house_veto_100]
house_v100_LCV
house_v100_name

#length(senate_veto_100)
#senate_veto_100
#length(house_veto_100)
#house_veto_100

```

Now, lets make some graphs to visually represent this distribution:
```{r}
ggplot(house_life, 
       aes(x = lifetime_score, 
           fill = party,
           color = party)) +
  geom_histogram(position = "identity", 
                 alpha = 0.5,
                 binwidth = 1) +
  scale_color_manual(values = c("dodgerblue2", "tomato2")) +
  scale_fill_manual(values = c("dodgerblue2", "tomato2")) +
  geom_vline(aes(xintercept = house_median),
             color = "black",
             linetype="solid")+
  theme_minimal()

ggplot(senate_life, 
       aes(x = lifetime_score, 
           fill = party,
           color = party)) +
  geom_histogram(position = "identity", 
                 alpha = 0.5,
                 binwidth = 1) +
  scale_color_manual(values = c("dodgerblue2", "tomato2")) +
  scale_fill_manual(values = c("dodgerblue2", "tomato2")) +
  geom_vline(aes(xintercept = senate_median),
             color = "black",
             linetype="solid")+
  theme_minimal()

```

```{r}
LCV_life_plot <- LCV_life %>% 
  group_by(chamber,
           party,
           lifetime_score) %>% 
  dplyr::summarise(count=n()) %>% 
  mutate(score_hou_sen = ifelse(chamber == 1,count,count*-1))

ggplot(LCV_life_plot, 
       aes(x = lifetime_score,
           y = score_hou_sen,
           fill = party,
           color = party)) +
  geom_bar(stat = "identity",
           alpha = 0.5) +
  scale_color_manual(values = c("dodgerblue2", "tomato2")) +
  scale_fill_manual(values = c("dodgerblue2", "tomato2")) +
  guides(color = FALSE,
         fill = FALSE)+
  geom_vline(aes(xintercept = house_median),
             color = "black", linetype="solid")+
  geom_text(aes(x=house_median, label="House Median", y=20),
            color="black",  angle=90, vjust = -0.75, size=3, family="serif")+
  geom_vline(aes(xintercept = senate_median),
             color = "black", linetype="solid")+
  geom_text(aes(x=senate_median, label="Senate Median", y=20),
            color="black",  angle=90, vjust = -0.75, size=3, family="serif")+
  geom_vline(aes(xintercept = house_veto_0),
             color = "black", linetype="solid")+
  geom_text(aes(x=house_veto_0, label="House Veto Pivot", y=20),
            color="black",  angle=90, vjust = -0.75, size=3, family="serif")+
   geom_vline(aes(xintercept = senate_veto_0),
             color = "black", linetype="solid")+
  geom_text(aes(x=senate_veto_0, label="Senate Veto Pivot", y=20),
            color="black",  angle=90, vjust = -0.25, size=2.5, family="serif")+
  geom_vline(aes(xintercept = senate_fili_L),
             color = "black", linetype="solid")+
  geom_text(aes(x=senate_fili_L, label="Filibuster Pivot Left", y=20),
            color="black",  angle=90, vjust = -0.75, size=3, family="serif")+
  geom_vline(aes(xintercept = senate_fili_R),
             color = "black", linetype="solid")+
  geom_text(aes(x=senate_fili_R, label="Filibuster Pivot Right", y=20),
            color="black",  angle=90, vjust = -0.75, size=3, family="serif")+
   geom_hline(yintercept=0)+
  xlim(100.5,-0.5)+
  scale_y_continuous(breaks = c(-10,0,10,20,30,40),
                       labels = abs(c(-10,0,10,20,30,40)))+
  geom_text(aes(x=50, label="Senate", y=-7),
            color="tan4", size=5, family="serif")+
  geom_text(aes(x=50, label="House", y=10),
            color="tan4", size=5, family="serif")+
  labs(title = "U.S. Congress LCV Scores",
       subtitle = "Lifetime League of Conservation Voters Scores",
       x = "LCV Score",
       y = "Count")+
  theme_minimal()+
  theme(text = element_text(family = "serif"),
    plot.title = element_text(size=18, face="bold"),
    plot.subtitle = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size=10, face="bold", vjust = -1),
    axis.title.y = element_text(size=10, face="bold", vjust = 2))

ggsave("us_LCV.png", width = 8, height = 5)

```

